---
title: "Exeter Automobile Receivables Trust - Case"
author: "Will Palmquist"
format: 
  pdf:
   fig-width: 10
   fig-height: 20
   documentclass: article
editor: visual
pdf-engine: pdflatex
header-includes:
   \usepackage{float}
   \floatplacement{table}{H}
---

## [ABS-EE Filing Data]{.underline}

The Exeter Automobile Receivables Trust 2022-6 is an asset backed security comprised of subprime auto loans. The monthly loan level reporting data has the following characteristics:

-   31635 auto loans

-   19 months of data from 2022-10-01 to present

### Accessing the Data

To access, download, and aggregate the monthly asset data files, see `01-SEC-Scraping.R` in the associated github repo found here:

## [Exploratory Analysis]{.underline}

The following figures identify relevant trends regarding the

```{r}
#| echo: false
1 + 1
```

You can add options to executable code like this

```{r}
#| echo: false
2 * 2
```

The `echo: false` option disables the printing of code (only output is displayed).

## Forecasting: Charged-Off Principal

## Kelley-Blue Book Values
